---
title: "Esports EDA"
author: "Sean Conner"
date: "`r Sys.Date()`"
output: 
  rmdformats::downcute:
    downcute_theme: "chaos"
    self_contained: true
    number_sections: true
    code_folding: show
    toc_float: true
    df_print: paged
---

# Setup

```{r setup, message=F, warning=F}

library(tidyverse)
library(glue)

```

```{css, echo=FALSE}

pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 400px;
}

```


# Load the data
```{r load_data}

esports_csv_load_list <-
  list.files(pattern = ".csv", full.names = T) %>%
  set_names(gsub("^\\./", "", .) %>%
              gsub(".csv$", "", .)) %>%
  map(data.table::fread)


general_data <- esports_csv_load_list$GeneralEsportData

historical_data <- esports_csv_load_list$HistoricalEsportData

```

# Helper function to deobfuscate potential join issues

```{r inner_join_w_diagnostics}
  
inner_join_w_diagnostics <-
  function(left_df, right_df, by = NULL) {
    df_list <- list(left_df, right_df)
    
    larger_df_index <- df_list %>%
      map_int(~ nrow(.x)) %>%
      order(decreasing = T) %>%
      head(1)
    
    larger_df <- df_list[[larger_df_index]]
    
    larger_df_records <- larger_df %>%
      nrow()
    
    inner_joined_df <- suppressMessages(df_list %>%
                                          reduce(inner_join, by = by))
    
    joined_records <- inner_joined_df %>%
      nrow()
    
    join_relative_discrep <- joined_records - larger_df_records
    join_relative_discrep_pretty_print <-
      scales::comma(join_relative_discrep,
                    prefix = ifelse(join_relative_discrep > 0, "+", ""))
    
    match_perc <- round(joined_records / larger_df_records, 6)
    
    match_perc_pretty_print <- paste0(match_perc * 100, "%")
    
    unmatched_perc <- round(abs(1 - match_perc), 6)
    unmatched_perc_pretty_print <- paste0(unmatched_perc * 100, "%")
    
    message_text <- case_when(
      join_relative_discrep == 0 ~ "Pre- & post-inner-join record counts match exactly! You may continue without issue.",
      join_relative_discrep > 0 ~ "More post- than pre-inner-join records. Investigate possible duplicates that increased the result set.",
      join_relative_discrep < 0 &
        join_relative_discrep > (larger_df_records * -1) ~
        "More pre- than post-inner-join records, but some matches exist. Investigate records lacking a match between the input datasets.",
      join_relative_discrep == (larger_df_records * -1) ~ "Warning: Zero matching records in the join result! Is it possible that join keys were misspecified?"
    )
    
    message(
      glue(
        "{message_text}

    Records difference: {join_relative_discrep_pretty_print}
    Matched %: {match_perc_pretty_print}
    Unmatched %: {unmatched_perc_pretty_print}"
      )
    )
    
    if (join_relative_discrep == 0) {
      return(inner_joined_df)
    }
    
    else {
      left_right_anti <- left_df %>%
        suppressMessages(anti_join(right_df))
      
      right_left_anti <- right_df %>%
        suppressMessages(anti_join(left_df))
      
      inner_left_anti <- inner_joined_df %>%
        suppressMessages(anti_join(left_df))
      
      inner_right_anti <- inner_joined_df %>%
        suppressMessages(anti_join(right_df))
      
      left_inner_anti <- left_df %>%
        suppressMessages(anti_join(inner_joined_df))
      
      right_inner_anti <- right_df %>%
        suppressMessages(anti_join(inner_joined_df))
      
      inner_join_group_dupes <- inner_joined_df %>%
        group_by(across(!where(is.numeric))) %>%
        count() %>%
        ungroup() %>%
        filter(n > 1) %>%
        arrange(desc(n))
      
      list(
        inner_joined_df = inner_joined_df,
        match_stats = data.frame(
          join_relative_discrep = join_relative_discrep,
          match_percent = match_perc,
          unmatched_percent = unmatched_perc
        ),
        inner_join_group_dupes = inner_join_group_dupes,
        left_right_anti = left_right_anti,
        right_left_anti = right_left_anti,
        inner_left_anti = inner_left_anti ,
        inner_right_anti = inner_right_anti,
        left_inner_anti = left_inner_anti ,
        right_inner_anti = right_inner_anti
      )
    }
    
  }

```

# Diagnose and resolve join issues

With a simple class test on `diag_join_issues` using `inherits()`, we can check whether our helper function returned an object of class "list", indicating there are some match stats to review and attempt to resolve, or a single inner-joined data frame with no detected issues. 

The below code will print the `match_stats` if there were join issues to review found by our helper function `inner_join_w_diagnostics` or a message stating it's safe to proceed otherwise. Since I've run this interactively (**spoiler alert**) before writing this, I know that there are diagnostics to review in this case.

So, let's take a look at `diag_join_issues$match_stats` and understand what that's telling us.

## Match stats
diag_join_issues$match_stats shows that there is a positive 1 record discrepancy in the join results compared to the largest of the two data frames in the join, meaning we got 1 **additional** record in our join results compared to what we expected.

Let's turn to `diag_join_issues$inner_join_group_dupes` now as our first look into why that might be.

```{r}

diag_join_issues <- inner_join_w_diagnostics(general_data, historical_data)

if(inherits(diag_join_issues, "list")){
  diag_join_issues$match_stats
} else {
  message("No match stats to report! Proceed without issue.")
}

```

## Group duplicates

```{r}

diag_join_issues$inner_join_group_dupes

```


```{r find_general_data_duplicate}

# general_data should have one row per game; find game names where there are more than one row

general_data_dupes <- general_data %>% 
  filter(Game %in% 
           (diag_join_issues$inner_join_group_dupes %>% 
           distinct(Game) %>% 
           pull(Game)))

general_data <- suppressMessages(general_data %>% 
  anti_join(general_data_dupes %>%
              slice(2)) %>% 
  mutate(ReleaseDate = fct_inseq(factor(ReleaseDate, ordered = T))))


# The general dataset has two records for the game Resistance 2.
#
# There are 4 numeric, non-grouping columns in the dataset: TotalEarnings,
# OnlineEarnings, TotalPlayers, and TotalTournaments

# ReleaseDate is also numeric, but it is a representation of the release year, 
# which is more appropriate to treat as a categorical variable in this context.)  
#
# One row for Resistance 2 has valid, non-zero values for 3 of the 4 numeric, 
# non-grouping variables in the dataset: TotalEarnings, TotalPlayers, and
# TotalTournaments. It has a value of 0 in the OnlineEarnings column.
#
# The other row for Resistance 2 shows 0 for all of the numeric, non-grouping 
# columns in the dataset: TotalEarnings, OnlineEarnings, TotalPlayers, and
# TotalTournaments.
#
# Since the Resistance 2 row showing 0 for all of the numeric, non-grouping
# columns does not provide any additional information and is seemingly simply an
# erroneous duplicate row, that row is dropped from the joined dataset via an
# anti-join. The ReleaseDate variable is also converted to a factor.


```

```{r}

final_join_data <- inner_join_w_diagnostics(general_data, historical_data)

```
 
```{r}

DT::datatable(final_join_data,
              filter = 'top',
              options = list(pageLength = 5))

```

```{r}

# Of course, let's skim our data and start to clearly understand what the
# dataset's characteristics are. From here, we can assess additional cleaning
# steps that might be necessary and start to make some hypotheses about the 
# data.

final_join_data %>% 
  skimr::skim()

```
 
## Points of Interest

* The Tournaments variable from the historical data is often not 1 for a given date. In fact, the average number of tournaments for a given date is ~6.5. However, the median is 2.0



```{r include = F, eval = F}
# possibly use forecast package to predict
fixed_general_data <- general_data %>% 
  filter(!(Game == "Resistance 2" & TotalEarnings == 0))

# Let's get rid of games w/ zero earnings and zero tournaments
# For everything else, we can do some imputation by converting zeros to NA,
# since, in many cases, there's nonsensical zeros (i.e. Earnings but no Tournaments and vice versa).
fixed_general_data %>% 
  arrange(-ReleaseDate) %>% 
  filter(!(TotalEarnings == 0 & TotalTournaments == 0))



# Let's see if we can predict a tournament's earnings based on the release date
# of the game, the genre of the game, and the number of players in the
# tournament.

fixed_general_data %>% 
  filter(TotalEarnings == 0)



fixed_general_data %>% 
  arrange(-TotalEarnings)

fixed_general_data %>% 
  lm(TotalEarnings ~ TotalPlayers + TotalTournaments + Genre, .) %>% 
  summary()

```


